---
title: "Verifying the 'Central Limit Theorem' for Exponential Distributions"
author: "Ronald Stalder"
date: '`r Sys.Date()`'
output:
  html_document:
    highlight: tango
    theme: journal
  pdf_document:
    highlight: tango
---

#### Data Science - Statistical Inference - Course Project - Part 1


### Sinopsis

The ***Central Limit Theorem (CLT)*** states that for any distribution of the population and given a large enough number of samples $n$, the distribution of these samples $\bar X$ is approximately normal and centered at the mean of the population, i.e. the mean $\mu_{\bar x} = \mu$. The standard deviation of $\bar X$, called *standard error*, is given by $SE(\bar X) = \frac{\sigma}{\sqrt n}$ where $\sigma$ is the standard deviation of the population.

The statements of the CLT are to be verified for the example of the "population" being the *exponential distribution*, which has mean and standard deviation $\mu = \sigma = \frac{1}{\lambda}$. $\lambda$ is the *scale parameter* of the distribution, which will be set to $\lambda = 0.2$.

The *"large enough number of samples"* depends on whether the distribution of the population is skewed (which is the case in our simulation here) and whether it contains outliers. As given by the assignment, this number will be set to $n = 1000$.

\pagebreak

### Generate the Simulation Data

```{r sim, results='hold'}
# set parameters (see sinopsis)
lambda <- 0.2     # scale parameter for exp. distribution
nsim <- 1000      # number of simulations
nsmpl <- 40       # number of samples for each simulation
n <- nsim * nsmpl # total number of samples
set.seed(516101)  # ensure reproducibility, see note below

# theoretical values to be compared against
Mu <- 1/lambda    # mean of exponential distribution
Var <- Mu^2       # variance; square of standard deviation (same as mean)
tSE <- Mu/sqrt(n) # theoretical standard error

# generate sample space Xbar for n simulations
Xbar <- Xbar.var <- Xbar.mean <- Xbar.SE <- Xbar.varm <- NULL # init vectors
for (i in 1:nsim) {                             # run {nsim} simulations
    simul <- rexp(nsmpl, lambda)                # generate {nsmpl} random samples
    Xbar <- c(Xbar, mean(simul))                # add their mean to Xbar
    Xbar.var <- c(Xbar.var, var(simul))         # add their variance
    # in order to show convergence, add "cummulative" values
    # (as if the simulation stopped here)
    Xbar.mean <- c(Xbar.mean, mean(Xbar))       # mean so far
    Xbar.SE <- c(Xbar.SE, ifelse(i == 1, 0,     # standard error so far, avoid NA
                          sd(Xbar)/sqrt(i)))
    Xbar.varm <- c(Xbar.varm, mean(Xbar.var))   # mean of simulation variances
}
SE <- sd(Xbar)/sqrt(nsim)                          # standard error

```


**note:**  
`set.seed(516101)` is used in order to ensure the reproducibility of the report. Not all seed values show the results equally well; various randomly picked values have been tested. The value of **516101** has been chosen for showing a good convergence within 1,000 simulations. For `set.seed(866376)`, for example, about 80,000 simulations would be needed.


\pagebreak

### Sample Mean versus Theoretical Mean

The CLT states that the mean of the samples should be approximately equal to the population mean, which is $\mu = \frac{1}{\lambda} = \frac{1}{`r lambda`} = `r 1/lambda`$. 

The mean of our sample space is $E[\bar X] =$ ` mean(Xbar) ` $= `r round(mean(Xbar),4)`$ (rounded to 4 decimals).   
This **confirms the CLT**; the difference is only `r abs(round(Mu - mean(Xbar),4))`.

The following plot shows how the sample mean converges to the population mean with increasing number of simulations:


```{r mean, results='hold'}
plot(Xbar.mean, type="l", xlab="number of simulations", ylab="mean of simulation means",
     main="E[Xbar] for increasing number of simulations", cex.main=0.7, cex.lab=0.8)
abline(h=Mu, col="blue")
legend("topright","population mean", lty = "solid", lwd = 2, col = "blue", bty = "n")
```



\pagebreak

### Sample Variance versus Theoretical Variance

The CLT states that the *standard error*, which is the standard deviation of the means of the samples, should be approximately $SE[\bar X] = \frac{\sigma}{\sqrt n}$, where $\sigma$ is the standard deviation of the population and $n$ is the *total number of samples* - in our case the number of simulations times the number of samples in each:  
Theoretical standard error $= \frac{`r Mu`}{\sqrt{`r nsim` \cdotp `r nsmpl`}} = `r tSE`$  

The standard error of our sample space is the standard deviation of $\bar X$ divided by the number of simulations: ` sd(Xbar)/sqrt(nsim) ` $=$ `r round(SE,4)` (rounded to 4 decimals). This **confirms the CLT**.  

Reformulating the statement about *standard error*, the CLT also says that the mean of the variances of the samples should be approximately equal to the population variance, which is $VAR(\bar X) = \sigma^2 = (\frac{1}{`r lambda`})^2 = `r (1/lambda)^2`$. 

The mean variance of our samples is `mean(Xbar.var)` $=$ `r round(mean(Xbar.var),4)` (rounded to 4 decimals). Again **confirming the CLT**.

The following two plots show the convergence for an increasing number of simulations:  
- left: $SE[\bar X]$ converges to $\frac{\sigma}{\sqrt n}$  
- right: mean of sample variance converges to population variance


```{r var, results='hold'}
par(mfrow = c(1,2))

plot(Xbar.SE, type="l", xlab="number of simulations", ylab="standard error of means",
     main="standard error", cex.main=0.7, cex.lab=0.8)
abline(h=tSE, col="blue")
legend("topright","theoretical SE", lty = "solid", lwd = 2, col = "blue", bty = "n")

plot(Xbar.varm, type="l", xlab="number of simulations", ylab="mean of sample variance",
     main="mean sample variance", cex.main=0.7, cex.lab=0.8)
abline(h=Var, col="blue")
legend("topright","population variance", lty = "solid", lwd = 2, col = "blue", bty = "n")
```


\pagebreak

### Normal Distribution of the Sample Space

It would be beyond the scope of this exercise (and the 6 page limit) to verify *analytically* that the distribution of $\bar X$ is indeed approximately normal . But the following two plots give a **clear visual confirmation**:  

- (left) the histogram of $\bar X$ overlayed by the curve of the theoretical normal distribution with the given mean and sd  
- (right) the *q-q plot* shows also that the quantiles for $\bar X$ fit nicely with the quantiles of the *standard* normal distribution.



```{r norm}
par(mfrow = c(1,2))
# Histyogram of Xbar overlayed with normal curve of same mean and sd
hist <- hist(Xbar, breaks=30, col="red", xlab="Simulation Means", 
          main=paste0("Histogram of Simulation Means\n",
                      "Overlayed with Normal"), cex.main=0.8)
# curve of theoretcial normal distribution
line_x <- seq(min(Xbar),max(Xbar),length.out=100)       # x coords of curve: 100 points
normal <- dnorm(x=line_x, mean=Mu, sd=Mu^2/sqrt(nsim))  # normal distribution densities
line_y <- normal * hist$counts[1] / hist$density[1]     # y: normal, streched to histogram
lines(line_x, line_y, col="blue", lwd=2)                # draw the curve
abline(v=Mu, col="grey")                                # vertical line at mean

# Q-Q plot comparing quantiles of Xbar versus standard normal distribution
qqnorm(Xbar, datax=T, col=rgb(1,0,0,0.2), pch=16, cex.main=0.8)
qqline(Xbar, datax=T, col="blue", lwd=2)
abline(h=0, col="grey")                                 # theoretical mean = 0
abline(v=Mu, col="grey")                                # Xbar mean = Mu
```

\pagebreak

### About This Document
  
The report was produced using `RStudio/knittr`  
**on `r format(Sys.time(), "%F at %H:%M:%S (%Z, GMT%z)")`**  

Numbers printed in the text part of this document are *not hardcoded*; they are the result of embedded R-code. The document can be generated with a different set of parameters without inconsistencies.

The R-Markdown source file has been published [here on Github](https://github.com/ronasta/RepData_PeerAssessment2)  

```{r sys-1, results='hold'}
if (Sys.which("uname") != "") system("uname -srpi", intern=TRUE)
if (Sys.which("lsb_release") != "") 
    print(paste("Ubuntu:",system("lsb_release -rc", intern=TRUE)[1]))
#print(paste("Rstudio version:", rstudio::versionInfo()$version))  # does not work in Rmd
```
```{r sys-2}
sessionInfo()
```
