---
title: "DatSci-06 statInf Quiz Week 1"
author: "Ronald Stalder"
date: '`r Sys.Date()`'
output:
  html_document:
    highlight: tango
    keep_md: yes
    theme: united
---

## Question 1

```
Consider influenza epidemics for two parent heterosexual families. Suppose that the probability is 17% that at least one of the parents has contracted the disease. The probability that the father has contracted influenza is 12% while the probability that both the mother and father have contracted the disease is 6%. What is the probability that the mother has contracted influenza?

(Hints look at lecture 2 around 5:30 and homework question on page 3/10).

5%
17%
11%
6%
```

Formula:  

$$ P(F\cup M) = P(F) + P(M) - P(FM) $$
$$ \Rightarrow P(M) = P(F\cup M) - P(F) + P(FM) $$


```{r q1}
pAny <- 0.17               # at least one
pF <- 0.12                 # father
pFM <- 0.06                # both

pM <- pAny - pF + pFM      # mother

print(paste0("Answer Q1: ", 100*pM, "%"))  ## 11%
```


## Question 2

```
A random variable, X is uniform, a box from 0 to 1 of height 1. (So that its density is f(x)=1 for 0≤x≤1.) What is its 75th percentile? 

(Hints, look at lecture 2 around 21:30 and homework 1 page 4/10. Also, look up the help function for the qunif command in R.)

0.75
0.10
0.25
0.50
```

- obviously it's **0.75**  
- let's check:  

```{r q2}
set.seed(123)             # make it reproducible
summary(runif(1000000))   # high number to get good 3. quantile (=75%)
```

- answer Q2: **1: 0.75**


## Question 3

```
You are playing a game with a friend where you flip a coin and if it comes up heads you give her X dollars and if it comes up tails she gives you Y dollars. The probability that the coin is heads is p (some number between 0 and 1.) What has to be true about X and Y to make so that both of your expected total earnings is 0. The game would then be called “fair”. 

(Hints, look at Lecture 4 from 0 to 6:50 and Homework 1 page 5/10. Also, for further reading on fair games and gambling, start with the Dutch Book problem ).

p/(1−p)=Y/X
p/(1−p)=X/Y
X=Y
p=X/Y
```

- The probalility for winning $X$ is $p$; for losing $Y$ is $(1-p)$  
- we want: $X\times p = Y\times(1-p)$  
- so: $\frac{p}{1-p} = \frac{Y}{X}$


- answer Q3: **1: p/(1-p)=Y/X**


## Question 4

```
A density that looks like a normal density (but may or may not be exactly normal) is exactly symmetric about zero. (Symmetric means if you flip it around zero it looks the same.) What is its median? 

(Hints, look at quantiles from Lecture 2 around 21:30 and the problem on page 9/10 from Homework 1.)

The median must be 0.
We can't conclude anything about the median.
The median must be 1.
The median must be different from the mean.
```

Distributions where *mean* $\neq$ *median* are skewed, for symmetrial curves they are the same.  

- answer Q4: **1: The median must be 0 (same as mean)**



## Question 5

```
Consider the following PMF shown below in R
x <- 1:4
p <- x/sum(x)
temp <- rbind(x, p)
rownames(temp) <- c("X", "Prob")
temp
##      [,1] [,2] [,3] [,4]
## X     1.0  2.0  3.0  4.0
## Prob  0.1  0.2  0.3  0.4

What is the mean? 

(Hint, watch Lecture 4 on expectations of PMFs and look at Homework 1 problem on page 10/10 for a similar problem calculating the variance.)

3
1
4
2
```

**Probability Mass Function PMF:**   

* on *discrete* random variables  
* must always be $\geq 0$  
* probabilities must sum up to $1$  


```{r q5}
x <- 1:4
p <- x/sum(x)
temp <- rbind(x, p)
rownames(temp) <- c("X", "Prob")
temp

# xp <- sum(x*p) or weighted.mean()

print(paste0("Answer Q5: ", weighted.mean(x, p)))   ## 3
```

## Question 6

```
A web site (www.medicine.ox.ac.uk/bandolier/band64/b64-7.html) for home pregnancy tests cites the following: “When the subjects using the test were women who collected and tested their own samples, the overall sensitivity was 75%. Specificity was also low, in the range 52% to 75%.” Assume the lower value for the specificity. Suppose a subject has a positive test and that 30% of women taking pregnancy tests are actually pregnant. What number is closest to the probability of pregnancy given the positive test?

(Hints, watch Lecture 3 at around 7 minutes for a similar example. Also, there's a lot of Bayes' rule problems and descriptions out there, for example here's one for HIV testing. Note, discussions of Bayes' rule can get pretty heady. So if it's new to you, stick to basic treatments of the problem. Also see Homework 2 question on page 5/12.)

10%
30%
20%
40%
```

for **one given** test:  

|----|----|----|----|  
|preg/test|positive|negative| |  
|true|TP|FN|P(TP)+P(FN)=1|  
|false|FP|TN| |  

- probability of *preg == true* is 30%
- **Sensitivity**: probability of true positive P(TP) = 75%  
- **Specificity**: probability of true negative P(TN) = 52%  
- P(FN) = 1 - p(TP) = 25%  
- P(FP) = 1 - p(TN) = 48%  

### solve using conditional probability: P(A given B)

$$P(A|B)=\frac{P(A\cap B)}{P(B)}$$  

where:

|----|-----|
|$A$ | is actually pregnant |  
|$P(A) = 0.3$ | 30% of women taking test are pregnant |  
|&nbsp;| |  
|$B$ | test is positive |  
|$P(B|A) = P(TP) = 0.75$ | positive test given pregnancy = sensitivity of test |  
|$P(A\cap B) = P(B|A)P(A)$ | probability of being pregnant AND positive test |  
| | see [Statistics Glossary](http://www.stats.gla.ac.uk/steps/glossary/probability.html#probability) |
|$\Rightarrow$|$=0.75\times 0.3$ = `r 0.75*0.3` |    
|&nbsp;| |  
|$P(B)$| positive test = true positives + false positives |  
| | (where woman is actually pregnant)|  
| | $= P(A)P(TP)+P(1-A)P(FP)$|  
| | $= P(A)P(TP)+P(1-A)P(1-P(TN))$ |  
|$\Rightarrow$| $= 0.3*0.75 + 0.7*(1-0.52)$ = `r 0.3*0.75 + 0.7*(1-0.52)` |  
|&nbsp;| |  
|$\frac{P(A\cap B)}{P(B)}$|= `r (0.75*0.3)/(0.3*0.75 + 0.7*(1-0.52))` |  

&nbsp;  

- answer Q6: **`r 10*round(10*(0.75*0.3)/(0.3*0.75 + 0.7*(1-0.52)))`%**
