---
title: "DatSci-06 statInf Quiz Week 2"
author: "Ronald Stalder"
date: "`r Sys.Date()`"
output: 
  html_document:
    highlight: tango
    keep_md: yes
    theme: united
---

## Question 1

```
What is the variance of the distribution of the average an IID draw of n observations from a population with mean μ and variance σ2.

σ/n
σ2/n
σ2
2σ/√n
```

Population: $$ Var(X) = E[(X - \mu)^2] = E[X^2] - E[X]^2 = \sigma^2 $$  

Sample: $$ S^2 = \frac{\sum_{i=1} (X_i - \bar X)^2}{n-1} $$  

- the sample variance is expected to be the population variance  

- answer Q1: **3: $\sigma^2$**  ## SHIT


### second try

- the sample *sd*, called *standard error* is $s = \frac{\sigma}{\sqrt n}$  
- the sample variation = $s^2 = \frac{\sigma^2}{n}$  

- answer Q1: **2: $\sigma^2 / n$**


## Question 2

```
Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?

16%
8%
22%
32%
```

```{r q2}
p <- pnorm(70, mean=80, sd=10, lower.tail = TRUE)

print(paste0("Answer Q2: ", round(100*p), "%"))  ## 16%
```

## Question 3

```
Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?

approximately 977
approximately 1175
approximately 1247
approximately 1223
```

```{r q3}
v <- qnorm(.95, mean=1100, sd=75)

print(paste0("Answer Q3: ", round(v)))  ## 1223
```

## Question 4

```
Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean?

approximately 1088 cc
approximately 1112 cc
approximately 1115 cc
approximately 1110 cc
```

- see [Wikipedia](http://en.wikipedia.org/wiki/Sampling_distribution)  
- the distribuition of the sample means is normal  
    - with mean = population mean $\mu$  
    - the *sd*, called *standard error* is $s = \frac{\sigma}{\sqrt n}$


```{r q4}
v <- qnorm(.95, mean=1100, sd=75/10)   # sd[pop] / sqrt(100)

print(paste0("Answer Q4: ", round(v)))  ## 1112
```

## Question 5

```
You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?

12%
3%
6%
19%
```

```{r q5}
p <- pbinom(4, size=5, prob=0.5, lower.tail=FALSE)  # 4+ successes, 5 tries, fair coin

print(paste0("Answer Q5: ", round(100*p), "%  SHIT"))  ## 3%  ## SHIT
```


### second try

- see 06_CommonDistros.Rmd Binominal Trials / Choose

```{r q5a, results='hold'}
p1 <- pbinom(3, size = 5, prob = .5, lower.tail = FALSE)  # MORE THAN 3 HEADS
p2 <- choose(5, 4) * .5 ^ 5 + choose(5, 5) * .5 ^ 5 
round(p1,4) == round(p2,4)
print(paste0("Answer Q5: ", round(100*p1), "%"))  ## 18%
```

## Question 6

```
The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?

47.5%
68%
95%
34%
```

- sample means tend to be normally ditributed, even if population is not
- sample mean = population mean = $\mu = 15$  
- sample *sd* = *standard error* is $s = \frac{\sigma}{\sqrt n} = 10 / \sqrt 100 = 1$
- p[14,16] = 1 - p[14-] - p[16+], or, because symmetric around mean: 1 - 2*p[14-]


```{r q6}

p <- 1 - 2 * pnorm(14, mean=15, sd=1, lower.tail=TRUE)

print(paste0("Answer Q6: ", round(100*p,1), "%"))  ## 68.3% => 68%
```

## Question 7

```
Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?

0.75
0.25
0.5
0.10
```

- the sample mean is expected to be the population mean  

- answer Q7: **3: 0.5**


## Question 8

```
The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?

0.03
0.08
0.06
0.12
```

* see *06_CommonDistros.Rmd line 221*   
* $X \sim Poisson(\lambda t)$ where 
  * $\lambda = E[X / t]$ is the expected count per unit of time
  * $t$ is the total monitoring time


```{r q8}
lambda = 5 * 3   # 5 people per hour for 3 hours
p <- ppois(10, lambda = lambda, lower.tail = TRUE)  # 10- people
  
print(paste0("Answer Q8: ", round(p,2)))  ## 0.12
```
